\chapter{Conclusions and future work}
\label{ch:conclusions}

This thesis investigated uplink non-orthogonal multiple access (NOMA) receivers and their impact on both link-level reliability and system-level resource sharing. 
The study first focused on the error-rate performance of different NOMA receivers in single-resource and multiple-resource systems, and then extended the analysis to coded multiple-resource and OFDM-based NOMA systems. Finally, a modulation coding scheme (MCS)-based user-pairing framework was developed to evaluate the system-level spectral efficiency (SE) behavior caused by NOMA resource sharing.

In Chapter~\ref{ch:single-resource}, single-resource uplink NOMA systems were studied. Power domain NOMA (PD-NOMA) with soft successive interference cancellation (SIC) and gain division multiple access (GDMA) were compared under small-scale Rayleigh block fading and different large-scale fading conditions. 
The results showed that soft SIC benefits from received-power disparity. However, when the number of users increases, soft SIC becomes more sensitive to error propagation, while GDMA provides stronger user-separation capability in the simulated single-resource cases.

In Chapter~\ref{ch:multi-resource}, the receiver analysis was extended to multiple-resource NOMA systems. MMSE-SIC detection was investigated with post-SINR ordering and amplitude-based ordering, and its BER performance was compared with HDS-CDMA with ML detection. 

In Chapter~\ref{ch:coded}, channel coding and iterative detection and decoding (IDD) structure were incorporated into the multiple-resource NOMA receiver. The BLER performance of MMSE-SIC-IDD, MMSE-PIC-IDD, and MAP-IDD was compared under different loading and modulation conditions. 
MAP-IDD generally provides strong decoding performance, especially under heavier loading, while MMSE-SIC-IDD can achieve competitive performance in moderate-loading cases.
The coded multiple-resource receiver framework was further extended to OFDM-based NOMA systems. The synchronous and asynchronous OFDM cases were considered. 
The simulation results demonstrate that MMSE-SIC-IDD with post-SINR ordering provides better robustness in several asynchronous scenarios.

In Chapter~\ref{ch:cqi}, the analysis moved from receiver-level performance to system-level resource sharing. 
The results show that NOMA user pairing does not necessarily increase the sum SE of the paired users, because the paired users usually suffer an MCS degradation. 
The system-level SE behavior is determined by the tradeoff between the paired users' SE loss and the SE contributed by the new users admitted through the released resources. 
The overall simulations also reveal that GDMA is more favorable than soft SIC for MCS-based pairing, since it's better receiver reliability and can support more effective after-pairing states and usually causes smaller SE loss.

In future research, several possible directions can be considered. First, the computational complexity of different NOMA receivers can be analyzed more quantitatively. A more detailed study of the number of candidate symbols, filtering operations, iterations, and implementation cost would provide a more complete performance-complexity tradeoff comparison.
Second, the MCS-based user-pairing analysis can be extended to multiple-resource NOMA receivers. 
Since multiple-resource NOMA receivers can improve the system performance, their pairing behavior may differ from the single-resource case. 
Extending the offline pairing table and system-level admission process to multiple-resource NOMA would provide a more general resource-sharing analysis.
Third, hybrid pairing methods may be investigated to ensure overall system-level SE increment. 